### 2017-06-12 mediation.Rmd
updated markdown script.

---
title: "Simple Mediation Example"
author: "Ryan Gan"
date: "6/7/2017"
output: html_document
---

```{r load libraries, warning=F, message=F}
library(broom) # for easy conversion of model results to dataframes
library(tidyverse) # data wrangling
library(MASS) # simulate correlations between two continous variables
```

Simulating a mediated dataframe where x -> m -> y

```{r sim mediated data}
# set df observations
n <- 500
# set seed
set.seed(303)

# simulate strong correlation between x and m

xm_data <- as_tibble(mvrnorm(n, mu = c(6, 4), Sigma = matrix(c(1,0.7,0.7,1), ncol = 2)))
# assign col names
colnames(xm_data) <- c("x", "m")
# check to amek sure correlation is ~70%
cor(xm_data)

# making continous x
sim_data <- xm_data %>% 
  # simulate protective effect between x and y
  mutate(y = rbinom(n, 1, 1/(1+exp(-(1+-0.6*m)))))
```

Some summary stats.

```{r outcome pr}
summary(sim_data)

prop.table(xtabs(~sim_data$y))
```

In our simulated data, our outcome makes up roughly 21% of our population. Thus OR will not approximate true risk.

First step is to establish the Total Effect (TE) relationship between x and y.

```{r xy association}
# check mod estimates
# this is referered to the direct effect
te_mod <- tidy((glm(y~x, sim_data, family="binomial"(link="logit"))))
te_mod

# going to start to fill in standard estimates and CIs
std_est <- as.data.frame(matrix(NA, ncol=4, nrow=3))
colnames(std_est) <- c("est", "rel_est", "lower_95", "upper_95")

# OR
std_est[1,1] <- "total_effect"
std_est[1,2] <- round(exp(te_mod[2,2]),3)
# 95 CI
std_est[1, 3:4] <- round(exp(c(te_mod[2,2]-1.96*te_mod[2,3],
                               te_mod[2,2]+1.96*te_mod[2,3])),3)
```

Second step is establishing x -> m, and m -> y relationship. These two estimates are required to meet the assumption, and to estimate the indirect/mediated path.

```{r xm and my association}
# linear x -> m mod beta
xm_mod <- tidy(glm(m~x, sim_data, family = "gaussian"(link="identity")))
xm_mod

# logit m -> y mod beta
my_mod <- tidy(glm(y~m, sim_data, family="binomial"(link="logit")))
my_mod

# OR
round(exp(my_mod[2,2]),3)
# 95 CI
round(exp(c(my_mod[2,2]-1.96*my_mod[2,3], my_mod[2,2]+1.96*my_mod[2,3])),3)
```

Now that each path of the DAG has been confirmed, we calculate the controled direct effect (CDE).

```{r cde}
cde_mod <- tidy(glm(y~ x + m, sim_data, family="binomial"(link="logit")))  
cde_mod

# OR
std_est[2,1] <- "cde"
std_est[2,2] <- round(exp(cde_mod[2,2]),3)
# 95 CI
std_est[2,3:4] <- round(exp(c(cde_mod[2,2]-1.96*cde_mod[2,3],
                              cde_mod[2,2]+1.96*cde_mod[2,3])),3)
```

Notice the association betwene x and y essentially is null when m is in the model. Some old epi text books say this is evidence of mediation, assuming m is an intermediate variable. This might be a decent hint, but we can actually quantify the indirect effect, sometimes refered to Natural Indirect Effect (NIE) when there is no interaction between the mediator and predictor.

When all our variables are continous TE - CDE = xm*teta2 = NIE

Where theta2 is m on y effect in the adjusted CDE model.

If all our variables are not continoous, then TE - CDE only approximates xm*theta2 = NIE. I'd advise use of the product method for most cases.

```{r nie estimates}
# te - cde
nie1 <- round(te_mod[2,2] - cde_mod[2,2], 3) 
# OR
exp(nie1)
cde_mod
# xm * my nie
nie2 <- round(xm_mod[2,2]*cde_mod[3,2],3)
# OR
exp(nie2)
```

General interpretation for the NIE:

A 1 unit increase in x as it passes through mediator m reduces the odds of y by 40%.

Next step is to estimate 95% confidence intervals. We have two choices, the product method/Sobel? test and bootstrapping. In this simulated example, they should be equivalent. For small sample sizes like with the 9health data, it's better to bootstrap as the Sobel test is likely unstable.

```{r nie se sobel}
# we will use estimates from these models to caculate the joint se
beta1 <- xm_mod[2,2]
theta2 <- cde_mod[3,2]
var_beta1 <- xm_mod[2,3]^2
var_theta2 <- cde_mod[3,3]^2

# calculate standard error of the nie
nie_se <- sqrt((theta2^2*var_beta1)+(beta1^2*var_theta2))

# estimate OR and 95% CI
std_est[3,1] <- "nie"
std_est[3,2:4] <- round(exp(c(nie2, nie2-1.96*nie_se, 
                                         nie2+1.96*nie_se)),3)

std_est
```

With small sample sizes, better to bootstrap bounds. I tend to use the percentile method since it's just easier to use with the broom package.

```{r mediation function}
# define function, mediation (simple function that doesn't have covariate)
mediation <- function(data, y, x, m, y_family, m_family){

 total_effect <- glm(y~x, data, family=y_family)$coefficients[2]
 cde <- glm(y~x+m, data, family=y_family)$coefficients[2]
 nie <- (glm(m~x, data, family=m_family)$coefficients[2]*
         glm(y~x+m, data, family=y_family)$coefficients[3])

 estimates <- cbind(total_effect, cde, nie)
 return(estimates)

}
```

Testing mediation function.

```{r testing mediation funciton}
# testing mediation function  
test <- mediation(sim_data,y="y", x="x", m="m", y_family="binomial",
                 m_family="gaussian")

test
```

Bootstrapped estimate terms. Median, 2.5 and 97.5 percentile should be good approximations of mean and 95% CI estimate terms.

```{r mediation boot}
# booting 1000 times 
boot_mediation <- sim_data %>% bootstrap(1000) %>% 
  do(tidy(mediation(data=., y="y", x="x", m="m", y_family="binomial",
                 m_family="gaussian")))

# estimates (empty matrix)
est <- matrix(NA, ncol=4, nrow=3) 
colnames(est) <- c("est", "median", "lower_95", "upper_95")

# fill total effect values
est[1,1] <- "total_effect"
est[1,2] <- round(exp(quantile(boot_mediation$total_effect, 0.5)),3)
est[1,3] <- round(exp(quantile(boot_mediation$total_effect, 0.025)),3)
est[1,4] <- round(exp(quantile(boot_mediation$total_effect, 0.975)),3)

# fill cde values
est[2,1] <- "cde"
est[2,2] <- round(exp(quantile(boot_mediation$cde, 0.5)),3)
est[2,3] <- round(exp(quantile(boot_mediation$cde, 0.025)),3)
est[2,4] <- round(exp(quantile(boot_mediation$cde, 0.975)),3)

# fill nie values
est[3,1] <- "nie"
est[3,2] <- round(exp(quantile(boot_mediation$nie, 0.5)),3)
est[3,3] <- round(exp(quantile(boot_mediation$nie, 0.025)),3)
est[3,4] <- round(exp(quantile(boot_mediation$nie, 0.975)),3)

booted_estimates <- as_data_frame(est)

# booted estimates
knitr::kable(booted_estimates, caption = "Booted Estimated Effects")

# standard estimates
knitr::kable(std_est, caption = "Standard Estimated Effects")
```