### 2017-06-20 categorical_msm.Rmd
updated

---
title: "Categorical MSM"
author: "Ryan_Gan"
date: "June 7, 2017"
output: html_document
---

```{r libraries, message=F}
library(tidyverse) 
library(broom) # for tidy model output and bootstrapping
library(MSModelR) # my package for simple msm functions
library(contrast)
```

## Simulated dataset

Simulating a simple continous distribution and a linear relationship with binary outcome which I'll then break in to an ordinal catogory by tertiles. No confounding yet.

```{r sim data}
n <- 500
set.seed(333)

# create normal distribution of x with mean of 16 and sd of 4
sim_data <- data_frame(x = rnorm(n, 16, 4)) %>% 
  # simulate y data with linear logit relationship to x
  mutate(y = rbinom(n, 1, 1/(1+exp(-(-1+0.10*x)))),
         # split in to three categories
         x_cat = ifelse(x < quantile(x,1/3), 0,
                 ifelse(x > quantile(x,2/3), 2, 1)),
         # dummy variable
         x0 = ifelse(x_cat == 0, 1, 0),
         x1 = ifelse(x_cat == 1, 1, 0),
         x2 = ifelse(x_cat == 2, 1, 0),
         x1_2 = ifelse(x_cat == 1 | x_cat == 2, 1, 0))
```

Truth.

```{r standard estimates}
# check density plots of x category given x
ggplot(sim_data, aes(x=x, fill=as.factor(x_cat))) +
  geom_density(alpha = 0.5)

# estimates good enough for example
true_mod <- glm(y~x, sim_data, family="binomial"(link="logit"))
tidy(true_mod)

# plot predicted probability of the linear funciton
pr_y_x <- predict(true_mod, sim_data, type="response")

# plot of predicted probability of continous variable
ggplot(sim_data, aes(x=x, y=pr_y_x)) +
  geom_point()

```

Standard analysis asking what the risk ratio is strata to strata

```{r risk ratio}
# 2x3 table
y_x <- xtabs(~x_cat+y, sim_data)
pr_yx <- prop.table(y_x, 1)
pr_yx
# difference in risk from strata 2 to 1
pr_yx[2,2] - pr_yx[1,2]
# difference in risk from strata 3 to 1
pr_yx[3,2] - pr_yx[1,2]
# difference in risk from strata 3 to 2
pr_yx[3,2] - pr_yx[2,2]

# estimate risk ratio
risk_ratio_mod <- glm(y~x1+x2, sim_data, family="binomial"(link="logit"))
tidy(risk_ratio_mod)

# comparing difference in x2 to x1
risk_ratio_mod2 <- glm(y~x2+x0, sim_data, family="binomial"(link="logit"))
tidy(risk_ratio_mod2)

# treating the categorical variables as continous
risk_ratio_mod3 <- glm(y~x_cat, sim_data, family="binomial"(link="logit"))
tidy(risk_ratio_mod3)

# now what if we combine groups x 3 and 2 together and compared against 1
risk_ratio_mod4 <- glm(y~x1_2, sim_data,family="binomial"(link="logit"))
tidy(risk_ratio_mod4)
```

Loss of information by combining the mid and high strata. But would this truly reflect a population effect where we may ask the question what if we reduced to background levels represented by strata 0? Something to think about.

Trying out MSM estimates.