### 2017-08-03 msm_introduction.Rmd
Finished msm presentation.

---
title: "Introduction to Marginal Structural Models"
author: "Ryan Gan"
date: "7/10/2017"
output: 
  slidy_presentation:
    incremental: true
---

## Setup

I will use the tidyverse and broom package for this presentation.

```{r setup, message=F, warning=F}
library(tidyverse)
library(broom)
```

## Causal inference

- Causal inference methods are growing in popularity (particularly among the epidemiology methodologists).

- Terms associated with these methods include: DAGs, counter factual, potential outcomes, g-estimation, g-computation, g-formula, g-something, inverse-probability-of-treatment weighting, marginal structural models, doubly-robust estimation, targeted maximum likelihood, etc.

- Jargon around this field makes these methods intimidating.

- This presentation will focus on marginal structural models (MSMs) and provide an introduction to two of the more common of these methods.

## What's the big deal about causal inference methods?

- Another tool/skill set to help evaluate causal relationships.

- Emphasis mostly placed on the analysis of data.

- Causal inference language is different from statistical language (i.e. X affecting Y vs. X correlated with Y) although we may run the exact same statistical test to assess.

## Three terms you'll frequently encounter

1. Directed Acyclic Graphs (DAGs): help to understand potential relationships between variables (i.e. potential collider and confounding bias)and how data may have been generated (i.e. temporal sequence exposure affects outcome).

2. Counter factual/potential outcomes: makes explicit the "what if?" scenario.

3. MSMs provide consistent estimates of differences or ratios of average (marginal) potential outcomes with less restrictions compared to standard regression methods (we'll return to this when we talk about time-dependent confounding).

## What is a counter factual?

- Randomized control trial: randomly assign exposure and no exposure to two groups and follow them over time to see which group is more likely to develop the outcome.

- What is implicit in this design is that the 'no exposure' group represents what would have happened to the 'exposure' group.

- A missing data problem: a subject's exposure and outcome can be observed only once.

- The counter factual solves our missing data problem and takes our "what if?" questions from implicit to explicit in our analysis of data.

## Marginal Structural Models (MSMs)

- Uses the principal of the counter factual and allows us to estimate an unbiased and consistent effect,under a less-restrictive set of assumptions:

1. Counter factual consistency

2. Exchangeability

3. Positivity

## Counterfactual consistency

- Allows us to equate observed outcomes for a given exposure value to the potential outcomes that would be observed under the same exposure value.

- This assumption depends on exposure assignment mechanism (i.e. RCT).

## Exchangeability

- Potential outcome under exposure a~0~ and a~1~ are independent of the observed exposures A~0~ and A~1~.

- Can be tested empirically.

- Easier to understand as being able to exchange exposure groups and observing the same distribution of the outcome.

- Conditional exchangeability happens when there is no unmeasured confounding.

## Positivity

- Probability of exposure is not 0 or 1.

- Can be empirically evaluated to make sure there are exposed and unexposed individuals within each confounder strata and prior exposure levels.

## Two commonly used MSMs

While there are more classes of MSMs we'll cover the most basic two

1. Parametric g-formula (sometimes called simple substitution)

2. Inverse-probability-of-treatment weighting (IPTW)

## Simulated data

I've simulated a simple example data frame with a known relationship to show you how to implement these two methods.

Simulate code is hidden, but here are the first 6 observations and a 2x2 table.
```{r simple simulation, echo=FALSE}
# set sample size and seed 
n <- 5000; set.seed(777)
# simulate exposure/treatment data with equal probability
sim_data <- as_tibble((rbinom(n, 1, prob = 0.5))) %>% 
  rename(A = value) %>% 
  # simulate protective effect between A and Y
  mutate(Y = rbinom(n, 1, 1/(1+exp(-(-2.5 + -0.6*A)))))
# print first 6 observations
head(sim_data)
#2x2 table
xtabs(~A + Y, sim_data)
```

## MSM using g-computation

- G-computation is a way in which we can estimate the counter factual state where everyone is treated/exposed (denoted by a = 1) and no one is treated/exposed (denoted by a = 0) and is implemented in a 3-step process:

1. Estimation of the conditional mean outcome E(Y|A).

2. Obtaining predicted potential outcomes under a = 1 and a = 0.

3. Estimate the statistical parameter (difference/ratio) by substituting the predicted mean outcomes under a=1 and a=0.

## Estimation of the conditional mean outcome E(Y|A)

This can be estimated with your standard generalized linear model. We'll estimate the risk ratio, but risk difference can be estimated as well.

```{r risk ratio}
# standard ratio model
ratio_model <- glm(Y~A, sim_data, family = "binomial"(link="logit"))
# ratio estimate
round(exp(ratio_model$coefficients[2]),2)
```
The treated group is 40% less likely to develop the outcome compared to the untreated group.

## Updating the predicted outcomes under a=1

First we duplicate our original data frame and then set everyone's exposure to 1.
```{r create a1 exposure dataframe}
# set everyone to exposed from our original dataframe
sim_data_a1 <- sim_data %>% mutate(A = 1)
# view first 6 records
head(sim_data_a1)
```
Now we apply our saved model coefficients (ratio_model) and updated the predicted values using the data frame where everyone receives exposure E(Y|a=1).
```{r pr a1}
# updated predicted probability of outcome 
pr_a1 <- predict(ratio_model, newdata = sim_data_a1, type = "response")
# average probability where everyone is exposed
mean(pr_a1)
```

## Updating the predicted outcomes under a=0

We repeat the same step under the scenario where no one is treated.
```{r create a0 exposure dataframe}
# set everyone to exposed from our original dataframe
sim_data_a0 <- sim_data %>% mutate(A = 0)
# view first 6 records
head(sim_data_a0)
```
Updating our predicted probability under no exposure scenario E(Y|a=0).
```{r pr a0}
# updated predicted probability of outcome 
pr_a0 <- predict(ratio_model, newdata = sim_data_a0, type = "response")
# average probability where no one is exposed
mean(pr_a0)
```

## Estimate the counterfactual parameter of interest

Now that we have a mean value for Y where everyone is treated and no one is treated, we can estimate our risk ratio.
```{r g forumla ratio}
# g formula ratio
round(mean(pr_a1)/mean(pr_a0),2)
```
Let's compare to our estimate using logistic regression again.
```{r standard est result}
# odds ratio (approximates risk ratio due to rare disease assumption)
round(exp(ratio_model$coefficients[2]),2)
```

We get pretty much the same result using the g-formula approach to a standard regression. However, the important thing is that we've now made our "what if?" question explicit and testable.

## MSM via IPTW

Another commonly use method is the inverse-probability-of-treatment weight.This is probably the most common marginal structural models (MSMs) methods used in practice today because it's easier to implement.

With IPTW estimation, confounding can be thought of as a problem of bias sampling where certain exposure-covariate subgroups are over or under represented compared to an randomized control trail. IPTW up-weights under-represented subjects from exposure-covariate strata.

## IPTW steps

1. Estimate the propensity score Pr(A|W)

2. Create weights

3. Estimate statistical parameter

## Estimation of the propensity of Pr(A)

Much like g-computation, our first step involves a standard regression model. However, unlike g-computation, the first step is to estimate the probability of treatment/exposure (A).

```{r iptw step 1}
# intercept only model since we don't have covariates
prop_mod <- glm(A ~ 1, family = "binomial", data = sim_data)
# summary of model
```

Easy enough.

## Estimating probability of a=1 and a=0

- For each observation, we estimate the probability of being exposed and simply take 1 minus that probability to calculate the probability of not being exposed.

- Before we create the weights, we first need to see if we have positivity, or experimental treatment assignment (ETA),violations (since these are simulated data and there are no factors that affect exposure probability, there won't be).

- Assessing ETA violations is easy enough to do by looking at a distribution of the predicted probabilities of A.

```{r iptw step 2.1}
# vector of propensity exposed for each observation
prob_exposed <- predict(prop_mod, type = "response")
# vector of propensity not exposed
prob_unexposed <- 1 - prob_exposed 
```

## Creating weights

- The weight (wt) is simply 1 over the probability of being exposed or unexposed. For those who were exposed, we assign wt as 1/pr(a=1). For unexposed, we assign wt as 1/pr(a=0).

```{r iptw step 2.2}
# create weights
wt <- as.numeric(sim_data$A == 1)/prob_exposed + 
  as.numeric(sim_data$A == 0)/prob_unexposed

# checking to make sure we inverted probabilities correctly
head(data.frame(sim_data$Y, sim_data$A, 1/prob_exposed, 1/prob_unexposed, wt))
```

## Estimating the statistical parameter

Now that we have the weights, we take all the subjects where A=1, multiply the weight and outcome and divide that by all the subjects where A=0 multiplied by the weights and outcome.

```{r iptw step 3}
# ratio estimate
iptw_ratio <- mean(as.numeric(sim_data$A==1)*wt*sim_data$Y)/
  mean(as.numeric(sim_data$A==0)*wt*sim_data$Y)
# estimate
round(iptw_ratio,2)
```

## What's the point?

- This is a lot of work to arrive at the same point estimate that can be estimated with a line of code.

- We haven't estimated any sort of inferential statistics like confidence intervals.

- G-computation and IPTW don't have empirical variance formulas (to my knowledge), therefore confidence intervals need to be estimated by bootstrapping or other sandwich approximations.

- There are two scenarios where these 'causal' estimates from MSMs will differ from standard regression approaches and where they are the most appropriate statistical approach

## When an MSM will provide a better estimate

1. The estimation of exposure on outcome is confounded by a time-dependent factor.

2. The estimation of exposure on outcome in the presence of interaction.

## MSM and a time-dependent confounder

- MSMs can help us get unbiased average/marginal estimates of exposure/treatment (A) on an outcome (Y) when A is administered at multiple time points and when a third factor (Z) is effected by A at time 0 and affects A at time 1.

- I'm going to use the example from Naimi et al. 2017, *An introduction to g methods*. It's a nicely written paper for those new to these causal inference methods and it also has SAS code for SAS users.

- In the example, we want to know the average effect of a binary treatment for HIV (A) on continuous CD4 counts (Y). Treatment is administered at multiple time points, with the only other covariate being a binary cutoff of HIV viral load (Z), where Z = 1 for viral load > 200 copies/ml. Further, viral load influences the decision to treat at that time point.

## Causal structure of our example

![](naimi_fig.png)

## Simulating example Naimi data

The Naimi paper uses the SAS cards statement to provide strata estimates. I've simulated a nearly identical data frame (code hidden due to size).

```{r time dependent Z sim data, echo=F}
# hiding simulation since it takes up a lot of space
# i'm sure there is a better way to simulate this, but eh
# simulating table 1 (kind of)
set.seed(123)
strata1 <- tibble(obs = 1:209271) %>%
  mutate(strata = 1, A0 = 0, Z1 = 0, A1 = 0, 
    # i'm simulating a little variance around the value provided by Naimi
    Y = rnorm(n = 209271, mean = 87.29, sd = 10))

strata2 <- tibble(obs = 1:93779) %>% 
  mutate(strata = 2, A0 = 0, Z1 = 0, A1 = 1, 
    Y = rnorm(n = 93779, mean = 112.11, sd = 12))

strata3 <- tibble(obs = 1:60654) %>%
  mutate(strata = 3, A0 = 0, Z1 = 1, A1 = 0, 
    Y = rnorm(n = 60654, mean = 119.65, sd = 16))

strata4 <- tibble(obs = 1:136293) %>%
  mutate(strata = 4, A0 = 0, Z1 = 1, A1 = 1, 
    Y = rnorm(n = 136293, mean = 144.84, sd = 10))

strata5 <- tibble(obs = 1:134781) %>%
  mutate(strata = 5, A0 = 1, Z1 = 0, A1 = 0, 
    Y = rnorm(n = 134781, mean = 105.28, sd = 14))

strata6 <- tibble(obs = 1:60789) %>%
  mutate(strata = 6, A0 = 1, Z1 = 0, A1 = 1, 
    Y = rnorm(n = 60789, mean = 130.18, sd = 11))

strata7 <- tibble(obs = 1:93903) %>%
  mutate(strata = 7, A0 = 1, Z1 = 1, A1 = 0, 
    Y = rnorm(n = 93903, mean = 137.72, sd = 14))

strata8 <- tibble(obs = 1:210527) %>%
  mutate(strata = 8, A0 = 1, Z1 = 1, A1 = 1, 
    Y = rnorm(n = 210527, mean = 162.83, sd = 9))

# bind all in to one dataset
td_sim_data <- rbind(strata1, strata2, strata3, strata4, strata5, strata6, 
                     strata7, strata8)
# remove to save room
rm(strata1, strata2, strata3, strata4, strata5, strata6, strata7, strata8)
```

```{r summary td data frame, echo=F}
summary_td_sim_data <- td_sim_data %>% group_by(strata) %>% 
  summarise(A0 = mean(A0), Z1 = mean(Z1), A1 = mean(A1), Y = mean(Y),
            N = n())

knitr::kable(summary_td_sim_data, 
  caption = "Simulated average outcome values by strata of treatment (A) and confounder (Z)")
```

## Testable research questions

There are a couple that come to mind:

1. What is the average level of CD4 counts in those that received treatment at time 0 compared to those never treated?

2. What is the average level of CD4 counts in those that received treatment at time 1 compared to those never treated?

3. What is the average level of CD4 counts in those treated at both time 0 and time 1 compared to those never treated?

## Standard estimates of the average effect of A on Y

We'd like to know the average effect of treatment (at both time points).

For reference, the 'true' average effect of A0 and A1 on Y is an increase of 50 CD4 count.

```{r time dependent std regression, echo = F}
std_est_table <- as_tibble(matrix(NA, nrow = 6, ncol = 2))
colnames(std_est_table) <- c("model_params","a_estimate")

# generate a vector of formulas to estimate and fill first column
std_est_table[,1] <- formula_vector <- c("Y~A0+A1+A0*A1", "Y~A0+A1+A0*A1+Z1", 
  "Y~A0", "Y~A0+Z1", "Y~A1", "Y~A1+Z1")

# for loop to estimate standard regression estimates
for(i in 1:6){
  x <- formula_vector[i]
  mod <- lm(as.formula(x), data=td_sim_data)
  std_est_table[i,2] <- round(sum(mod$coefficients[
    # subset only A0, A1, or A0*A1
    names(mod$coefficients)=="A0" | names(mod$coefficients)=="A1" |
    names(mod$coefficients)=="A0*A1"]),1) 
}

knitr::kable(std_est_table, 
  caption = "Standard regression estimates of average effect of A on Y")
```

## Reframing the question

- None of our various models using standard linear regression yielded the true value of treatment effect on CD4 count of 50.

- Let's ask the question what would the effect of treatment be on CD4 count if everyone were treated at both time points vs if no one was treated at both time points?

- MSM's will let us estimate this question.

## G-formula accounting for time-dependent confounder.

Our question that we wish to answer with our MSM can be written as follows:

$$
\psi = E(Y^{a_0=1,a_1=1} - Y^{a_0=0,a_1=0})
$$

We can ask other "what if?" questions like "what if everyone was only treated at time 0 vs not treated at other time points?" etc., but we'll just ask this question for now.

## Step 1: estimate standard relationships

- Estimate exposures (A~0~,A~1~) relationship with the outcome (Y).
- Estimate the effect of treatment (A~0~) on viral load (Z~1~).
- Estimate the effect of viral load (Z~1~) on treatment (A~1~).

```{r estimate associations}
# g formula
y_mod <- glm(Y ~ A1 + Z1 + A0 + A1*A0, data = td_sim_data, family = "gaussian")
z1_mod <- glm(Z1 ~ A0, data = td_sim_data, family = "binomial"(link="logit"))
a1_mod <- glm(A1 ~ Z1, data = td_sim_data, family = "binomial"(link="logit")) 
```

## Estimate Y where everyone is treated at both time points

- We first estimate $$P(Z_1=z_1=1 | A_0 = a_0=1)$$

- Followed by simulating a vector of new counter factual estimates of z~1~ values based on the prior probability of z~1~ = 1, given a~0~ = 1

```{r z1 update}
# fully exposed ----
z1_pr <- mean(predict(z1_mod, newdata = mutate(td_sim_data, A0 = 1), 
                      type = "response"))
# simulate vector of new Z1 values accounting for prior A0
z1_vec <- ifelse(rbernoulli(n = nrow(td_sim_data), z1_pr)==T,1,0)
```

## Alternate step

- We can also estimate counter factual estimates of a~1~ if we wanted to ask "what if?" questions like "what if we only treated at time 1?"

- But since we are asking treatment at both time points, this step is not needed.
```{r a1 update}
# update a1 
a1_pr <- mean(predict(a1_mod, newdata = mutate(td_sim_data, Z1 = z1_vec), 
                      type = "response"))
# simulate new a1 given z1 under a0 estimate
a1_vec <- ifelse(rbernoulli(n = nrow(td_sim_data), a1_pr)==T,1,0)
```

## Estimate Y under all exposed scenario.

- Update our Z1 with z~1~ value in scenario where a~0~ = 1.
- Updated our A1 and A0 value to all be 1.
- If you had a new a~1~ vector, that could be added as well.

```{r all exposed estimate}
# update predicted Y where everyone is exposed at time 0 and 1, and accounting for Z1 | A0 =1 
all_exposed <- mean(predict(y_mod, newdata = mutate(td_sim_data, 
                A1 = 1, A0 = 1, Z1 = z1_vec), type = "response"))
all_exposed
```

## Repeat the same process for everyone being unexposed
```{r unexposed estimate}
# unexposed
z1_pr <- mean(predict(z1_mod, newdata = mutate(td_sim_data, A0 = 0), 
          type = "response"))
z1_vec <- ifelse(rbernoulli(n = nrow(td_sim_data), z1_pr)==T,1,0)
all_unexposed <- mean(predict(y_mod, newdata = mutate(td_sim_data, 
                  A1 = 0, A0 = 0, Z1 = z1_vec), type = "response"))
all_unexposed
```

## Estimate the difference in scenarios

We can now calculate our estimated average effect of treatment on Y where everyone is treated and no one is treated.

```{r g comp difference}
# diff
g_diff <- all_exposed - all_unexposed
g_diff
```

```{r g comp td formula, eval = F, echo = F}
# hiding this for now, but general formula's will be needed if you want to bootstrap
# creating a formula. will need dplyr
g.formula.td <- function(data, y, a0, a1, z1, y.model, y.family,
  a1.model, a1.family, z1.model, z1.family, estimate = "diff"){

  un_exp_df <- exp_df <- data
  un_exp_df[,c(a0,a1)] <- 0
  exp_df[, c(a0,a1)] <- 1

  # define y.model
  q_mod <- glm(as.formula(paste(y, y.model, sep = "~")), data = data, 
    family = y.family)
  # define z1.model
  z1_mod <- glm(as.formula(paste(z1, z1.model, sep = "~")), data = data, 
    family = z1.family)
  # define a1.model
  a1_mod <- glm(as.formula(paste(a1, a1.model, sep = "~")), data = data, 
              family = a1.family) 
  # updated probability of z1 given a0 = 1
  z1_exp_pr <- mean(predict(z1_mod, newdata = exp_df, type = "response"))
  # simulate vector of new Z1 values accounting for prior A0 where everyone is exposed
  z1_exp_vec <- as.numeric(ifelse(rbinom(n = nrow(data),size = 1 , z1_exp_pr)==T,1,0))
  # update z value under a0=1 scenario
  exp_df[,z1] <- z1_exp_vec
  # update y estimate as if everyone was exposed
  y_exp <- mean(predict(y_mod,newdata=exp_df,type = "response"))

  # updated probability of z1 given a0 = 0
  z1_unexp_pr <- mean(predict(z1_mod, newdata = un_exp_df, type = "response"))
  # simualte vector of new Z1 values accounting for prior A0 where everyone is unexposed
  z1_unexp_vec <- as.numeric(ifelse(rbinom(n = nrow(data), size = 1, z1_unexp_pr)==T,1,0))
  un_exp_df[,z1] <- z1_unexp_vec
  # udpate y estimates as if everyone was unexposed
  y_unexp <- mean(predict(y_mod,newdata=un_exp_df,type = "response"))

  if(estimate == "diff"){ 
    msm_estimate <- y_exp - y_unexp
    }
  if(estimate == "ratio"){
    msm_estimate <- y_exp/y_unexp
  }
  return(msm_estimate)
  }

g.formula.td(data = td_sim_data, y = "Y", a0 = "A0", a1= "A1", z1 = "Z1",
  y.model = "A0+A1+Z1+A0*A1", y.family = "gaussian", a1.model = "Z1", 
  a1.family = "binomial", z1.model = "A0", z1.family = "binomial", estimate = "diff")

```

## IPTW with a time-dependent confounder

First we estimate the propensity of A~0~, A~1~, and A~1~|Z~1~ using standard regression models

```{r iptw td}
# define propensity models (intercept only model without covariates)
pr_a0_mod <- glm(A0 ~ 1, data = td_sim_data, family="binomial")
# probability of being treated at time 1
pr_a0 <- predict(pr_a0_mod, type = "response")
# define propensity a1 model
pr_a1_mod <- glm(A1 ~ 1, data = td_sim_data, family = "binomial")
# probability of being treated at time 2
pr_a1 <- predict(pr_a1_mod, type = "response")
# probability of a1 given z1
pr_a1_z_mod <- glm(A1 ~ Z1, data = td_sim_data, family = "binomial")
# probability of a1 given z1
pr_a1_z <- predict(pr_a1_z_mod, type = "response")
```

## Estimating weights

```{r iptw weigts td}
# create weights in dataframe
td_sim_data <- td_sim_data %>% 
  mutate(wt = ifelse(A0 == 1 & A1 == 1, (pr_a0/pr_a0)*(pr_a1/pr_a1_z),
    ifelse(A0 == 1 & A1 == 0, (pr_a0/pr_a0)*((1-pr_a1)/(1-pr_a1_z)),
    ifelse(A0 == 0 & A1 == 1, ((1-pr_a0)/(1-pr_a0))*(pr_a1/pr_a1_z),
    ifelse(A0 == 0 & A1 == 0, ((1-pr_a0)/(1-pr_a0))*((1-pr_a1)/(1-pr_a1_z)),NA
    )))))

# table of summarized weights
wt_table <- td_sim_data %>% 
  group_by(strata) %>% 
  summarise(A0 = mean(A0), Z1 = mean(Z1), A1 = mean(A1), Y = mean(Y), n = n(),
            wt = mean(wt), pseudo_n = n*wt)

head(wt_table)
```

## Run a generalized model with the weights in the model

We see we get the same result of an average effect of treatment on CD4 outcome is 50.

```{r}
# weighted model
iptw_mod <- glm(Y ~ A0 + A1 + A0*A1, data = td_sim_data, weights = wt,
                family = "gaussian")
# calculate difference
iptw_diff <- sum(iptw_mod$coefficients[names(iptw_mod$coefficients)=="A0" | 
  names(iptw_mod$coefficients)=="A1" | names(iptw_mod$coefficients)=="A0*A1"],1)
# difference
iptw_diff
```

## Comparison with our standard estimates

We can see that both MSM terms give us our 'true' value, where our standard regression terms are off by a bit.

Whether you think this is a big enough deal to use an MSM is up to you :)

```{r std est w msm est, echo = F}
std_est_msm_table <- std_est_table %>% 
  add_row( model_params = c("G-Formula", "IPTW"), 
           a_estimate = c(g_diff, iptw_diff))

knitr::kable(std_est_msm_table, 
  caption = "Standard and MSM estimates")
```

## MSMs when there is interaction with exposure

Using an example out of **Implementation of G-Compuation on a Simulated Data Set: Demonstration of a Causal Inference Technique** by Snowden et al. 2011.

This is example is probably a more practical reason to use MSMs to estimate a population-level effect.

The example simulated an effect of a binary cutoff of high ozone (A) on FEV (Y), a continuous measure of pulmonary function. It also include sex (W1) as a confounding variable and an effect modifier for a binary race (W2) that has no base term. This means that race does not effect FEV, but that exposure is modified by race, attenuating the effect on FEV.

The problem here is that we want to estimate a population-level effect for ozone on FEV, regardless of race.

```{r simulate data, echo =F}
# sample size
n <- 300
set.seed(285)
simdata <- data.frame(W1 = rbinom(n, 1, 0.4), W2 = rbinom(n, 1, 0.5))
# add simulated data of A
simdata <- transform(simdata, A = rbinom(n, 1, (0.5+0.2*W1-0.3*W2)))
# add simulated data of Y
simdata <- transform(simdata, Y = rnorm(n, (3-0.5*A+W1+0.3*A*W2), 0.4))
```

## Standard estimation approach

```{r snowden standard model}
reg_mod <- glm(Y~A+W1+A:W2, data = simdata, family = gaussian)
reg_mod
```

- If we want to know the causal effect of ozone (A) on FEV (Y), we have to some how account for the interaction with race (W2).

- For our reference race, high level of ozone decrease FEV by 0.48.

- For other race the effect of ozone decrease FEV by 0.18.

- To estimate a population-level effect, perhaps we could average the two effects?

- (0.48+0.18)/2 = 0.33

## Let's try our MSM methods

```{r msm with boot estimates, echo = F}
# percentile method with broom
# g formula
set.seed(321)
boot_msm <- simdata %>% bootstrap(1000) %>% 
  do(tidy(MSModelR::g.formula(data = ., y = "Y", a = "A", q.model = "A+W1+A:W2",
            model.family = "gaussian", estimate = "diff")))
# iptw
set.seed(123)
boot_iptw <- simdata %>% bootstrap(1000) %>% 
  do(tidy(MSModelR::iptw(data = ., y="Y",a="A",w="W1+W2",estimate="diff")))

# create empty matrix
estimates <- matrix(nrow = 2, ncol = 4, byrow = T)
colnames(estimates) <- c("msm_estimate", "boot_median", "lower_95", "upper_95")

# fill matrix for g formula 
estimates[1,1] <- "G-Formula"
estimates[1,2] <- round(quantile(boot_msm$x, 0.5),2)
estimates[1,3] <- round(quantile(boot_msm$x, 0.025),2)
estimates[1,4] <- round(quantile(boot_msm$x, 0.975),2)

# fill matrix for iptw
estimates[2,1] <- "IPTW"
estimates[2,2] <- round(quantile(boot_iptw$x, 0.5),2)
estimates[2,3] <- round(quantile(boot_iptw$x, 0.025),2)
estimates[2,4] <- round(quantile(boot_iptw$x, 0.975),2)

# convert matrix to dataframe
est_df <- data.frame(estimates)

knitr::kable(est_df, caption = paste0("MSM estimates with bootstrapped 95%CI"))
```

- G-formula estimate gives us the average value for each races of ozone on FEV.

- IPTW is a little over this average estimate, but different from anyone race estimate alone.

## MSMs

Cons:

- Producing estimates that may be interpreted as casual still requires assumptions like temporarily and no unmeasured confounding.

- May be a lot of effort (and computationally expensive) to arrive at the same answer you'd get with a standard regression.

- G-formula and IPTW don't handle continuous variables.

Pros:

- Most appropriate analytic method in some scenarios (time-dependent confounding).

- Going through the whole process of figuring out causal structures (DAGs), checking assumptions, and implementing MSMs provides an excellent framework and tool for assessing causal relationships.

## Cutting edge stuff

- Doubly-robust estimation: combines both the q-model and propensity model and if even one of those models is specified correctly the estimate will be correct.

- Targeted maximum likelihood (TMLE): uses an ensemble of machine learning algorithms to optimize the q-model and propensity model to control for confounders.

- Structural nested models: old method that can handle continuous exposures but was difficult to implement; new packages for R are making this more accessible to average users.