### 2021-08-10 propensity_example.Rmd

---
title: "Propensity Example"
author: "Ryan Gan"
date: "7/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Propensity Example

Following Lucy D'Agostino McGowan's blog and dissertation work to understand newer weight methods for propensity scores.

https://livefreeordichotomize.com/2019/01/17/understanding-propensity-score-weighting/

Simulate data.

```{r}
set.seed(928)
n <- 1000
X <- mvtnorm::rmvnorm(n,
                      mean = c(0.5, 1),
                      sigma = matrix(c(2, 1, 1, 1), ncol = 2)
)

dat <- tibble(
  x_1 = X[, 1],
  x_2 = X[, 2],
  treatment = as.numeric(- 0.5 + 0.25 * x_1 + 0.75 * x_2 + rnorm(n, 0, 1) > 0),
  outcome = 2 * treatment + rnorm(n, 0, 1)
)
head(dat)
```

Fit propensity.

```{r}
dat <- dat %>%
  mutate(
    propensity_score = glm(treatment ~ x_1 + x_2, data = dat, family = "binomial") %>%
      predict(type = "response")
  )
```

Estimate weights from propensity.

```{r}
dat <- dat %>%
  mutate(
    w_ate = (treatment / propensity_score) + 
      ((1 - treatment) / (1 - propensity_score)),
    w_att = ((propensity_score * treatment) / propensity_score) + 
      ((propensity_score * (1 - treatment)) / (1 - propensity_score)),
    w_atc = (((1 - propensity_score) * treatment) / propensity_score) + 
      (((1 - propensity_score) * (1 - treatment)) / (1 - propensity_score)),
    w_atm = pmin(propensity_score, 1 - propensity_score) / 
      (treatment * propensity_score + (1 - treatment) * (1 - propensity_score)),
    w_ato = (1 - propensity_score) * treatment + 
      propensity_score * (1 - treatment)
  )
```

View propensity in mirrored display.

```{r}
d <- dat %>%
  tidyr::spread(treatment, propensity_score, sep = "_p")

ggplot(d) + 
  geom_histogram(bins = 50, aes(treatment_p1)) + 
  geom_histogram(bins = 50, aes(x = treatment_p0, y = -..count..)) + 
  ylab("count") + xlab("p") +
  geom_hline(yintercept = 0, lwd = 0.5) +
  scale_y_continuous(label = abs) 
```

```{r}
ggplot(d) +
  geom_histogram(bins = 50, aes(treatment_p1), alpha = 0.5) + 
  geom_histogram(bins = 50, aes(treatment_p1, weight = w_ate), fill = "green", alpha = 0.5) + 
  geom_histogram(bins = 50, alpha = 0.5, aes(x = treatment_p0, y = -..count..)) + 
  geom_histogram(bins = 50, aes(x = treatment_p0, weight = w_ate, y = -..count..), fill = "blue", alpha = 0.5) + 
  ylab("count") + xlab("p") +
  geom_hline(yintercept = 0, lwd = 0.5) +
  scale_y_continuous(label = abs) 
```