### 2021-08-10 superlearner_vignette.Rmd

---
title: "SuperLearner Vignette"
author: "EC Polley"
date: "April 14, 2017"
output: html_document
---

## Purpose

Running through the SuperLearner Vignette to learn a bit about ensemble methods.

```{r libraries, echo = F}
library(SuperLearner)
library(glmnet)
library(kernlab)
library(KernelKnn)
```

Load the Boston housing data from MASS package. Seperating outcome from covariates. Creating training and confirmation datasets.

```{r boston data, echo = F}
# Setup example dataset.

# Load a dataset from the MASS package.
data(Boston, package = "MASS")

# Review info on the Boston dataset.
head(Boston)

# Check for any missing data - looks like we don't have any.
colSums(is.na(Boston))

# Extract our outcome variable from the dataframe.
outcome = Boston$medv

# Create a dataframe to contain our explanatory variables.
data = subset(Boston, select = -medv)

# Check structure of our dataframe.
str(data)

# If we had factor variables we would use model.matrix() to convert to numerics.

# Review our dimensions.
dim(data)

# Set a seed for reproducibility in this random sampling.
set.seed(1)

# Reduce to a dataset of 150 observations to speed up model fitting.
train_obs = sample(nrow(data), 150)

# X is our training sample.
X_train = data[train_obs, ]

# Create a holdout set for evaluating model performance.
# Note: cross-validation is even better than a single holdout sample.
X_holdout = data[-train_obs, ]

# Create a binary outcome variable: towns in which median home value is > 22,000.
outcome_bin = as.numeric(outcome > 22)

Y_train = outcome_bin[train_obs]
Y_holdout = outcome_bin[-train_obs]

# Review the outcome variable distribution.
table(Y_train, useNA = "ifany")
```

Load SuperLearner package and check available models. For maximum accuracy one might try at least the following models: glmnet, randomForest, XGBoost, SVM, and bartMachine. These should ideally be tested with multiple hyperparameter settings for each algorithm.

```{r sl library}
library(SuperLearner)
library(randomForest)
# review available models
listWrappers()

# look at glmnet model code
SL.glmnet

```

# Fit individual models

Let's fit 2 separate models: lasso (sparse, penalized OLS) and randomForest. We specify family = binomial() because we are predicting a binary outcome, aka classification. With a continuous outcome we would specify family = gaussian().

```{r rf and lasso models}
# Set the seed for reproducibility.
set.seed(1)

# Fit lasso model.
sl_lasso = SuperLearner(Y = Y_train, X = X_train, family = binomial(),
                        SL.library = "SL.glmnet")
sl_lasso

# Review the elements in the SuperLearner object.
names(sl_lasso)

# Here is the risk of the best model (discrete SuperLearner winner).
sl_lasso$cvRisk[which.min(sl_lasso$cvRisk)]

# Here is the raw glmnet result object:
str(sl_lasso$fitLibrary$SL.glmnet_All$object, max.level = 1)

# Fit random forest.
sl_rf = SuperLearner(Y = Y_train, X = X_train, family = binomial(),
                     SL.library = "SL.randomForest")
sl_rf
```

Risk is a measure of model accuracy or performance. We want our models to minimize the estimated risk, which means the model is making the fewest mistakes in its prediction. It's basically the mean-squared error in a regression model, but you can customize it if you want.

SuperLearner is using cross-validation to estimate the risk on future data. By default it uses 10 folds; use the cvControl argument to customize.

The coefficient column tells us the weight or importance of each individual learner in the overall ensemble. By default the weights are always greater than or equal to 0 and sum to 1. In this case we only have one algorithm so the coefficient has to be 1. If a coefficient is 0 it means that the algorithm isn't being used in the SuperLearner ensemble.

### Fit multiple models

Instead of fitting the models separately and looking at the performance (lowest risk), we can fit them simultaneously. SuperLearner will then tell us which one is best (discrete winner) and also create a weighted average of multiple models.

We include the mean of Y ("SL.mean") as a benchmark algorithm. It is a very simple prediction so the more complex algorithms should do better than the sample mean. We hope to see that it isn't the best single algorithm (discrete winner) and has a low weight in the weighted-average ensemble. If it is the best algorithm something has likely gone wrong.

```{r rf lasso and mean}
set.seed(1)
sl = SuperLearner(Y = Y_train, X = X_train, family = binomial(),
  SL.library = c("SL.mean", "SL.glmnet", "SL.randomForest"))
sl

# Review how long it took to run the SuperLearner:
sl$times$everything
```

Again, the coefficient is how much weight SuperLearner puts on that model in the weighted-average. So if coefficient = 0 it means that model is not used at all. Here we see that random forest is given the most weight, following by lasso.

### Predict on new data

Now that we have an ensemble let's predict back on our holdout dataset and review the results.
So we have an automatic ensemble of multiple learners based on the cross-validated performance of those learners, nice!

```{r predict on holdout}
# Predict back on the holdout dataset.
# onlySL is set to TRUE so we don't fit algorithms that had weight = 0, saving computation.
pred = predict(sl, X_holdout, onlySL = T)

# Check the structure of this prediction object.
str(pred)

# Review the columns of $library.predict.
summary(pred$library.predict)

# Histogram of our predicted values.
library(ggplot2)
qplot(pred$pred[, 1]) + theme_minimal()

# Scatterplot of original values (0, 1) and predicted values.
# Ideally we would use jitter or slight transparency to deal with overlap.
qplot(Y_holdout, pred$pred[, 1]) + theme_minimal()

# Review AUC - Area Under Curve
pred_rocr = ROCR::prediction(pred$pred, Y_holdout)
auc = ROCR::performance(pred_rocr, measure = "auc", x.measure = "cutoff")@y.values[[1]]
auc

```

AUC can range from 0.5 (no better than chance) to 1.0 (perfect). So at 0.96 we are looking pretty good!

### Customize a model hyperparameter

Hyperparameters are the configuration settings for an algorithm. OLS has no hyperparameters but essentially every other algorithm does.

There are two ways to customize a hyperparameter: make a new learner function, or use create.Learner().

Let's make a variant of random forest that fits more trees, which may increase our accuracy and can't hurt it (outside of small random variation).

```{r sl mod of rf}
# Review the function argument defaults at the top.
SL.randomForest

# Create a new function that changes just the ntree argument.
# (We could do this in a single line.)
# "..." means "all other arguments that were sent to the function"
SL.rf.better = function(...) {
  SL.randomForest(..., ntree = 3000)
}

set.seed(1)

# Fit the CV.SuperLearner.
# We use V = 3 to save computation time; for a real analysis use V = 10 or 20.
cv_sl = CV.SuperLearner(Y = Y_train, X = X_train, family = binomial(), V = 3,
                        SL.library = c("SL.mean", "SL.glmnet", "SL.rf.better", "SL.randomForest"))

# Review results.
summary(cv_sl)
```

Looks like our new RF is not improving performance. This implies that the original 500 trees had already reached the performance plateau - a maximum accuracy that RF can achieve unless other settings are changed (e.g. max nodes).

For comparison we can do the same hyperparameter customization with create.Learner().

```{r custom learner}
# Customize the defaults for randomForest.
learners = create.Learner("SL.randomForest", params = list(ntree = 3000))

# Look at the object.
learners

# List the functions that were created
learners$names

# Review the code that was automatically generated for the function.
# Notice that it's exactly the same as the function we made manually.
SL.randomForest_1

set.seed(1)

# Fit the CV.SuperLearner.
# We use V = 3 to save computation time; for a real analysis use V = 10 or 20.
cv_sl = CV.SuperLearner(Y = Y_train, X = X_train, family = binomial(), V = 3,
                        SL.library = c("SL.mean", "SL.glmnet", learners$names, "SL.randomForest"))

# Review results.
summary(cv_sl)
```

We get exactly the same results between the two methods of creating a custom learner.

### Test algorithm with multiple hyperparameter settings

The performance of an algorithm varies based on its hyperparamters, which again are its configuration settings. Some algorithms may not vary much, and others might have far better or worse performance for certain settings. Often we focus our attention on 1 or 2 hyperparameters for a given algorithm because they are the most important ones.

For randomForest there are two particularly important hyperparameters: mtry and maximum leaf nodes. Mtry is how many features are randomly chosen within each decision tree node - in other words, each time the tree considers making a split. Maximum leaf nodes controls how complex each tree can get.

Let's try 3 different mtry options.

```{r hyperparameter settings}
# sqrt(p) is the default value of mtry for classification.
floor(sqrt(ncol(X_train)))

# Let's try 3 multiplies of this default: 0.5, 1, and 2.
mtry_seq = floor(sqrt(ncol(X_train)) * c(0.5, 1, 2))
mtry_seq

learners = create.Learner("SL.randomForest", tune = list(mtry = mtry_seq))

# Review the resulting object
learners

# Check code for the learners that were created.
SL.randomForest_1
SL.randomForest_2
SL.randomForest_3

set.seed(1)

# Fit the CV.SuperLearner.
# We use V = 3 to save computation time; for a real analysis use V = 10 or 20.
cv_sl = CV.SuperLearner(Y = Y_train, X = X_train, family = binomial(), V = 3,
                        SL.library = c("SL.mean", "SL.glmnet", learners$names, "SL.randomForest"))

# Review results.
summary(cv_sl)
```

We see here that mtry = 3 performed a little bit better than mtry = 1 or mtry = 7, although the difference is not significant. If we used more data and more cross-validation folds we might see more drastic differences. A higher mtry does better when a small percentage of variables are predictive of the outcome, because it gives each tree a better chance of finding a useful variable.

Note that SL.randomForest and SL.randomForest_2 have the same settings, and their performance is very similar - statistically a tie. It's not exactly equivalent due to random variation in the two forests.

A key difference with SuperLearner over caret or other frameworks is that we are not trying to choose the single best hyperparameter or model. Instead, we usually want the best weighted average. So we are including all of the different settings in our SuperLearner, and we may choose a weighted average that includes the same model multiple times but with different settings. That can give us better performance than choosing only the single best settings for a given algorithm, which has some random noise in any case.

#### Stopping here for now; the rest is about parallel computing