### 2021-10-29 causal_infernece_examples_v2.Rmd
update

---
title: "Causal inference methods examples part 2"
author: "Ryan Gan"
date: "2021-10-19"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 2
---

```{r setup}
# loading r packages 
library(tmle) # targeted maximum likelihood
library(MASS)
library(xgboost) # gradient boosting with l2 regularization
library(glmnet) # lasso, ridge, and elastic regularization
library(gbm) # gradient boosting machiens
library(shapr) # explainable tree modles
library(tidyverse) # general data wrangling and plotting
library(mgcv) # for non-linear terms
```

Start by simulating a relationship between the treatment and a binary outcome. 
Adding two known confounders that affect both treatment and outcome and adding
some random normal variables with no association with either treatment or outcome (although may happen by chance).

```{r sim}
# binary confounder example 
set.seed(100)
c1_x_association <- log(5)
c1_y_association <- log(5)
c2_x_association <- log(0.25)
c2_y_association <- log(0.25)

trt_out_association_true <- log(0.5) # true effect, protective effect

# sample size (note bias shrinks as sample size increases)
n <- 5000
# known confounders
c1 <- rbinom(n, 1, 0.6)
c2 <- rbinom(n, 1, 0.2)
age <- rnorm(n, mean = 55, sd = 5) # for non-linear function

# non-linear term on older patients less likely to get treatment

# 10 random noise variables to add to the model
rv_mat <- matrix(rnorm(n*10), ncol=10)
# give matrix some variable names 
colnames( rv_mat ) <- paste0('rv_', 1:10)

# relationship between confounders and treatment
trt <- rbinom(n, 1, (1 / (1+exp(-(c1*c1_x_association + c2*c2_x_association + 0.2*age + -0.0045*age^2)))))

# likelihood fun that includes relationship with treatment

lik_fun <- (
  1 / (1 + exp( -(trt*trt_out_association_true + c1*c1_x_association + c2*c2_x_association + 0.01*age ) ) )
  )

# binomial outcome based on the likelihood function
outcome <- rbinom(n, 1, lik_fun)

# put it all together in a dataframe to be used later
df <- data.frame( cbind(outcome, trt, c1, c2, age, rv_mat) )

```

2 x 2 and proportion table by treatment and outcome. In crude analysis, looks 
like both the relative risk and odds ratio suggest treatment increases likelihood
for the outcome. Note since our proportion of the outcome is high, the odds ratio
will not approximate the relative risk.

```{r 2x2}
tab <- xtabs(~ trt + outcome)
print(tab)

# crude relative risk
( tab[2,2] / sum(tab[2,]) ) / ( tab[1,2] / sum(tab[1, ]) )
# crude odds ratio
( tab[2,2] / tab[2,1])/ ( tab[1,2] / tab[1, 1] ) 
```

Estimation of crude associations with generalized linear models.
```{r crud_mods}
crude_trt_est <- broom::tidy( glm(outcome ~ trt, family = 'binomial') )[2, ]$estimate

# bias
print(paste0('Bias of True - Expected: ', round( crude_trt_est - trt_out_association_true , 2) ))
```

If we know the confounders of interest (based on previous studies or expert input),
a regression model works well enough for most cases.

```{r adj_mod}

# adjusted model formula with known confounders
adj_mod <- glm(outcome ~ trt + c1 + c2 + age, family = binomial(link='logit'), data = df)

# bias
print(
  paste0('Bias of True - Expected: ', 
         round( trt_out_association_true - adj_mod$coefficients['trt'], 2)
         )
  )
```

# G-computation

G-computation (a.k.a simple substitution, counterfactual estimate) first proposed by Robins in the epi literature sometime in the 1990s. In my opinion, this is Pearl's Do-calculus. Both state that if we specify a model `outcome | treatment, covariates` sufficiently to account for confounding bias, then we have achieved conditional exchangeability. We can then estimate the marginal effect of treatment on the outcome by taking the model and predicting as if everyone were treated, and predicting as if no one was treated. The average predicted score where everyone is treated can then be compared to the average predicted score where no one was treated and that should be the marginal risk difference (in a binary outcome) or relative 6ratio difference.

Using our adjusted model above, we can estimate this as follows:

```{r g_comp}
# copy dataframes
all_trt_df <- no_trt_df <- df
# set trt to all trt or no trt
all_trt_df$trt <- 1
no_trt_df$trt <- 0

# estimate using fit model and take the average of predictions
all_trt_pred <- predict(adj_mod, newdata = all_trt_df, type = 'response') 
no_trt_pred <- predict(adj_mod, newdata = no_trt_df, type = 'response')

# means for risk
trt_mean <- mean( all_trt_pred )
no_trt_mean <- mean( no_trt_pred )

# log odds ratio
mean_log_odds <- mean( log( ( all_trt_pred / (1 - all_trt_pred ) ) / (no_trt_pred / (1 - no_trt_pred)) ) )

# absolute reduction in risk
print(paste0( 'Risk difference: ', round( trt_mean - no_trt_mean , 2 ) ) )
# relative reduction in risk
print(paste0( 'Relative risk: ', round( trt_mean / no_trt_mean , 2 ) ) )
# odds ratio
gcomp_or <- ( trt_mean / ( 1 - trt_mean ) ) / (no_trt_mean / ( 1 - no_trt_mean ) )
print(paste0('Odds ratio: ', round( gcomp_or , 2)) )

# bias
print(
  paste0('Bias of True - Expected: ', 
         round( trt_out_association_true - mean_log_odds, 2)
         )
  )
```

All covs in model; saturated model.

```{r g_comp_glm_satorrect}
saturated_mod <- glm(outcome ~ trt + c1 + c2 + age + rv_1 + rv_2 + rv_3 + rv_4
                     + rv_5 + rv_6 + rv_7 + rv_8 + rv_9 + rv_10, 
                     family = binomial(link='logit'), data = df)

# estimate using fit model and take the average of predictions
all_trt_pred_sat <- predict(saturated_mod, newdata = all_trt_df, type = 'response') 
no_trt_pred_sat <- predict(saturated_mod, newdata = no_trt_df, type = 'response')

# means for risk
trt_mean_sat <- mean( all_trt_pred_sat )
no_trt_mean_sat <- mean( no_trt_pred_sat )

# log odds ratio
mean_log_odds_sat <- mean( 
  log( ( all_trt_pred_sat / (1 - all_trt_pred_sat ) ) / (no_trt_pred_sat / (1 - no_trt_pred_sat)) )
  )

# absolute reduction in risk
print(paste0( 'Risk difference: ', round( trt_mean_sat - no_trt_mean_sat , 2 ) ) )
# relative reduction in risk
print(paste0( 'Relative risk: ', round( trt_mean_sat / no_trt_mean_sat , 2 ) ) )
# odds ratio
gcomp_or_sat <- ( trt_mean_sat / ( 1 - trt_mean_sat ) ) / (no_trt_mean_sat / ( 1 - no_trt_mean_sat ) )
print(paste0('Odds ratio: ', round( gcomp_or_sat , 2)) )

# bias
print(
  paste0('Bias of True - Expected: ', 
         round( trt_out_association_true - mean_log_odds_sat, 2)
         )
  )

```

G-computation using lasso.

```{r g_comp_lasso}
# lasso
lasso_g_mod <- cv.glmnet(
  x = as.matrix( df[, 2:ncol(df)] ) , 
  y = df$outcome , 
  alpha = 1.0, # lasso penalty
  family = 'binomial'
)

# estimate using fit model and take the average of predictions
all_trt_pred_lasso <- predict(
  lasso_g_mod , 
  newx = as.matrix( all_trt_df[, 2:ncol(all_trt_df)] ), 
  s = "lambda.min", type = 'response'
)

no_trt_pred_lasso <- predict(
  lasso_g_mod , 
  newx = as.matrix( no_trt_df[, 2:ncol(no_trt_df)] ), 
  s = "lambda.min", type = 'response'
  )

# means for risk
trt_mean_lasso <- mean( all_trt_pred_lasso )
no_trt_mean_lasso <- mean( no_trt_pred_lasso )

# log odds ratio
mean_log_odds_lasso <- mean( 
  log( 
    ( all_trt_pred_lasso / (1 - all_trt_pred_lasso ) ) / (no_trt_pred_lasso / (1 - no_trt_pred_lasso))
    ) 
  )

# absolute reduction in risk
print(paste0( 'Risk difference: ', round( trt_mean_lasso - no_trt_mean_lasso , 2 ) ) )
# relative reduction in risk
print(paste0( 'Relative risk: ', round( trt_mean_lasso / no_trt_mean_lasso , 2 ) ) )
# odds ratio
gcomp_or_lasso <- ( ( trt_mean_lasso) / ( 1 - trt_mean_lasso ) ) / (no_trt_mean_lasso / ( 1 - no_trt_mean_lasso ) ) 
print(paste0('Odds ratio: ', round( gcomp_or_lasso , 2)) )

# bias
print(
  paste0('Bias of True - Expected: ', 
         round( trt_out_association_true - mean_log_odds_lasso, 2)
         )
  )

```

G-computation using xgboost.

```{r gcomp_xgb}
# xgb model
xgb_g_mod <- xgboost(
  data = as.matrix( df[, 2:ncol(df)] ),
  label = df$outcome, 
  max_depth = 4, eta = 0.01, nrounds = 5000, verbose = 0, nthread = 4, 
  objective = "binary:logistic"
)

# estimate using fit model and take the average of predictions
all_trt_pred_xgb <- predict(xgb_g_mod , newdata = as.matrix( all_trt_df[, 2:ncol(all_trt_df)] )) 
no_trt_pred_xgb <- predict(xgb_g_mod , newdata = as.matrix( no_trt_df[, 2:ncol(no_trt_df)] ))

# means for risk
trt_mean_xgb <- mean( all_trt_pred_xgb )
no_trt_mean_xgb <- mean( no_trt_pred_xgb )

# log odds ratio
mean_log_odds_xgb <- mean( 
  log( 
    ( all_trt_pred_xgb / (1 - all_trt_pred_xgb ) ) / (no_trt_pred_xgb / (1 - no_trt_pred_xgb))
    ) 
  )

# absolute reduction in risk
print(paste0( 'Risk difference: ', round( trt_mean_xgb - no_trt_mean_xgb , 2 ) ) )
# relative reduction in risk
print(paste0( 'Relative risk: ', round( trt_mean_xgb / no_trt_mean_xgb , 2 ) ) )
# odds ratio
gcomp_or_xgb <- ( ( trt_mean_xgb) / ( 1 - trt_mean_xgb ) ) / (no_trt_mean_xgb / ( 1 - no_trt_mean_xgb ) ) 
print(paste0('Odds ratio: ', round( gcomp_or_xgb , 2)) )

# bias
print(
  paste0('Bias of True - Expected: ', 
         round( trt_out_association_true - mean_log_odds_xgb, 2)
         )
  )

```

Plot showing bias from true.

```{r gcomp_bias_plot}
est <- c('Unadjusted GLM', 'Correct Adj GLM', 'G-Comp Adj', 'G-Comp All Cov', 
         'G-Comp Lasso', 'G-Comp XGB')

g_coef <- c(crude_trt_est, adj_mod$coefficients['trt'], log( gcomp_or ), 
            log( gcomp_or_sat), log( gcomp_or_lasso ), log( gcomp_or_xgb ) )

g_df <- data.frame(
  estimation  = 'G-Computation', 
  method = est , 
  coef = g_coef
  ) %>% 
  mutate(
    method = factor( method , levels = method ), 
    odds_ratio = exp( coef )
  )

ggplot(g_df, aes( x = method, y = odds_ratio )) +
  geom_point() +
  geom_hline(yintercept = exp( trt_out_association_true ), color = 'red') +
  ylim( 0.25, 1.25 ) +
  theme_linedraw() +
  ylab('Odds Ratio') +
  xlab('Estimation Method')
```

## Propensity score

Another way to estimate is via propensity score. With this approach we model
the likelihood of being treated. Once we have a model, we can then get the 
predicted probability of being treated given the covariates and create weights.
Inverse predicted probability (IPTW) is common, but there are other weighting methods
that can account for some biases the arise with IPTW. Once we have a predicted
probability, we can create weights. I'm using a couple models below to illustrate
how this works using a naive linear model where the age effect is mis-specified,
a GAM model where a penalized spine is used for age, and an xgboost tree model
using all predictors (as if we knew nothing about the variables).

Note, we have a non-linear effect of age on treatment that should be accounted for.
If age is linear, the propensity might not fit as well. I'm not too worried about
overfitting because I want to try and predict as best I can on the observed
population and I'm not interested in predicting on an unseen population.

```{r propensity_models}
# age as a linear term
lin_pr_mod <- glm(
  as.formula(
    paste0('trt ~ ', paste(colnames(df)[3:ncol(df)], collapse = ' + ' ) )
    ), family = 'binomial', data = df
  )

# known gam model
gam_pr_mod <- gam(
    as.formula(
    paste0('trt ~ s(age) + ', paste(colnames(df)[c(3,4,6:ncol(df))], collapse = ' + ' ) )
    ), 
  family = 'binomial', data = df
  )

# lasso
lasso_pr_mod <- cv.glmnet(
  x = as.matrix( df[, 3:ncol(df)] ) , 
  y = df$trt , 
  alpha = 1.0, # lasso penalty
  family = 'binomial'
)

# xgb model
xgb_pr_mod <- xgboost(
  data = as.matrix( df[, 3:ncol(df)] ),
  label = df$trt, 
  max_depth = 4, eta = 0.01, nrounds = 1000, verbose = 0, nthread = 4, 
  objective = "binary:logistic"
)

# get predictions from models back
lin_pred <- predict(lin_pr_mod, newdata = df, type = 'response')

gam_pred <- predict(gam_pr_mod, newdata = df, type = 'response')

lasso_pred <- predict(lasso_pr_mod, newx = as.matrix( df[, 3:ncol(df)] ), s = "lambda.min", type = 'response')
colnames(lasso_pred) <- 'lasso_pred'

xgb_pred <- predict(xgb_pr_mod, newdata = as.matrix( df[, 3:ncol(df)] ))

```

Applying different weights. Average treatment effect (ATE), 
average treatment effect among treated (ATT), average treatment effect among
evenly match (ATM), average treatment effect of observed (ATO).

```{r create_ate_weights}
p_weights <- function(outcome, treatment, propensity_score) {
  # ate
  w_ate = (treatment / propensity_score) + ((1 - treatment) / (1 - propensity_score))

  # att
  w_att = ((propensity_score * treatment) / propensity_score) + 
      ((propensity_score * (1 - treatment)) / (1 - propensity_score))

  # atc
  w_atc = (((1 - propensity_score) * treatment) / propensity_score) + 
      (((1 - propensity_score) * (1 - treatment)) / (1 - propensity_score))

  # atm
  w_atm = pmin(propensity_score, 1 - propensity_score) / 
      (treatment * propensity_score + (1 - treatment) * (1 - propensity_score))

  # ato
  w_ato = (1 - propensity_score) * treatment + propensity_score * (1 - treatment)

  # dataframe with weights
  weight_df <- data.frame(
    outcome = outcome, 
    treatment = treatment,
    propensity = propensity_score, 
    w_ate = w_ate,
    w_att = w_att,
    w_atc = w_atc, 
    w_atm = w_atm, 
    w_ato = w_ato
    )

  return(weight_df)
}

# weight lists
lin_wt_df <- p_weights(outcome = df$outcome, treatment = df$trt, propensity_score = lin_pred)
gam_wt_df <- p_weights(outcome = df$outcome, treatment = df$trt, propensity_score = gam_pred)
lasso_wt_df <- p_weights(outcome = df$outcome, treatment = df$trt, propensity_score = as.vector( lasso_pred ) )
xgb_wt_df <- p_weights(outcome = df$outcome, treatment = df$trt, propensity_score = xgb_pred)

```

Plots of propensity scores weighting schemes.

```{r wt_plots}
d <- lin_wt_df %>% 
  tidyr::spread(treatment, propensity, sep = "_p")

ggplot(d) +
  geom_histogram(bins = 50, aes(treatment_p1), alpha = 0.5) + 
  geom_histogram(bins = 50, aes(treatment_p1, weight = w_ate, fill = "Treated"), alpha = 0.5) + 
  geom_histogram(bins = 50, alpha = 0.5, aes(x = treatment_p0, y = -..count..)) + 
  geom_histogram(bins = 50, aes(x = treatment_p0, weight = w_ate, y = -..count.., fill = "Not Treated"),
                 alpha = 0.5) + 
  scale_fill_manual('', values = c('blue', 'red') ) +
  ylab("Count") + 
  xlab("Propensity") +
  ggtitle('ATE weighting of propensity: linear model') +
  geom_hline(yintercept = 0, lwd = 0.5) +
  scale_y_continuous(label = abs)  +
  theme_linedraw()

```

```{r lin_att}
ggplot(d) +
  geom_histogram(bins = 50, aes(treatment_p1), alpha = 0.5) + 
  geom_histogram(bins = 50, aes(treatment_p1, weight = w_att, fill = "Treated"), alpha = 0.5) + 
  geom_histogram(bins = 50, alpha = 0.5, aes(x = treatment_p0, y = -..count..)) + 
  geom_histogram(bins = 50, aes(x = treatment_p0, weight = w_att, y = -..count.., fill = "Not Treated"),
                 alpha = 0.5) + 
  scale_fill_manual('', values = c('blue', 'red') ) +
  ylab("Count") + 
  xlab("Propensity") +
  ggtitle('ATT weighting of propensity: linear model') +
  geom_hline(yintercept = 0, lwd = 0.5) +
  scale_y_continuous(label = abs)  +
  theme_linedraw()
```

```{r lin_atm}
ggplot(d) +
  geom_histogram(bins = 50, aes(treatment_p1), alpha = 0.5) + 
  geom_histogram(bins = 50, aes(treatment_p1, weight = w_atm, fill = "Treated"), alpha = 0.5) + 
  geom_histogram(bins = 50, alpha = 0.5, aes(x = treatment_p0, y = -..count..)) + 
  geom_histogram(bins = 50, aes(x = treatment_p0, weight = w_atm, y = -..count.., fill = "Not Treated"),
                 alpha = 0.5) + 
  scale_fill_manual('', values = c('blue', 'red') ) +
  ylab("Count") + 
  xlab("Propensity") +
  ggtitle('ATM weighting of propensity: linear model') +
  geom_hline(yintercept = 0, lwd = 0.5) +
  scale_y_continuous(label = abs)  +
  theme_linedraw()
```

Propensity models with weights.

```{r prop_wt_estimates}
# ate estimates
lin_prop_mod_ate <- glm(outcome ~ trt, data = lin_wt_df, family = 'quasibinomial', weights = w_ate) 
gam_prop_mod_ate <- glm(outcome ~ trt, data = gam_wt_df, family = 'quasibinomial', weights = w_ate) 
lasso_prop_mod_ate <- glm(outcome ~ trt, data = lasso_wt_df, family = 'quasibinomial', weights = w_ate) 
xgb_prop_mod_ate <- glm(outcome ~ trt, data = xgb_wt_df, family = 'quasibinomial', weights = w_ate)

# atm estimates
lin_prop_mod_atm <- glm(outcome ~ trt, data = lin_wt_df, family = 'quasibinomial', weights = w_atm) 
gam_prop_mod_atm <- glm(outcome ~ trt, data = gam_wt_df, family = 'quasibinomial', weights = w_atm) 
lasso_prop_mod_atm <- glm(outcome ~ trt, data = lasso_wt_df, family = 'quasibinomial', weights = w_atm) 
xgb_prop_mod_atm <- glm(outcome ~ trt, data = xgb_wt_df, family = 'quasibinomial', weights = w_atm) 
```

```{r propensity_estimates}
p_est <- c(
  'P-Score Lin ATE',
  'P-Score GAM ATE' , 
  'P-Score Lasso ATE',
  'P-Score XGB ATE' , 
  'P-Score Lin ATM',
  'P-Score GAM ATM' , 
  'P-Score Lasso ATM',
  'P-Score XGB ATM' 
  )

# propensity coef
p_coef <- c(
  # ate
  coef(lin_prop_mod_ate)['trt']
  , coef(gam_prop_mod_ate)['trt']
  , coef(lasso_prop_mod_ate)['trt']
  , coef(xgb_prop_mod_ate)['trt']
  # atm
  , coef(lin_prop_mod_atm)['trt']
  , coef(gam_prop_mod_atm)['trt']
  , coef(lasso_prop_mod_atm)['trt']
  , coef(xgb_prop_mod_atm)['trt'] 
  )

p_score_df <- data.frame(
  estimation = 'Propensity', 
  method = p_est, 
  coef = p_coef
  )  %>% 
  mutate(
    method = factor( method , levels = method ), 
    odds_ratio = exp( coef ), 
    percent_bias = coef - log(0.5)
  )

```

Plot for propensity score estimates difference from the true estimate.

```{r pscore_plot}
ggplot(p_score_df, aes( x = method, y = odds_ratio )) +
  geom_point() +
  geom_hline(yintercept = exp( trt_out_association_true ), color = 'red') +
  ylim( 0.25, 1.25 ) +
  theme_linedraw() +
  ylab('Odds Ratio') +
  xlab('Estimation Method') +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.5))
```

## Targeted-maximum likelihood estimation (TMLE)

This is a doubly-robust approach to estimating marginal effect that can use
machine learning models in an ensemble model.

```{r tmle}

# libraries to fit g and q formula in tmle
sl_libs <- c('SL.mean', 'SL.glm', 'SL.step', 'SL.glmnet', 'SL.gbm', 'SL.xgboost')

tmle_mod <- tmle(
  outcome, trt, as.matrix( df[, 3:ncol(df)] ) , family = 'binomial', 
  g.SL.library = sl_libs , 
  Q.SL.library = sl_libs , 
  V = 3, # use fewer cross validation folds, default it 10
)

```

```{r tmle_summary}

summary(tmle_mod)

```

```{r combind_metrics}
final_df <- g_df %>% 
  rbind(p_score_df) %>% 
  rbind(
    data.frame(
      estimation = 'TMLE',
      method = 'TMLE', 
      coef = -0.52012 , 
      odds_ratio = 0.59445
      )
    ) %>% 
  mutate(
    estimation = case_when(
      method %in% c('Unadjusted GLM', 'Correct Adj GLM') ~ 'Conditional Models', 
      TRUE ~ estimation
    )
  )

final_df
```

Final plot summarizing bias of estimates compared to true value.

```{r summary_plot}
ggplot(
  data = final_df, 
  aes( x = factor(method, levels = method), y = odds_ratio , color = estimation) ) +
  geom_point() +
  coord_flip() +
  scale_x_discrete(
    limits = rev(factor(final_df$method, levels = final_df$method))
    )  +
  geom_hline(yintercept = exp( trt_out_association_true ), color = 'red') +
  ylim( 0.25, 1.25 ) +
  theme_linedraw() +
  ylab('Odds Ratio') +
  xlab('Estimation Method')
```